
\documentclass{sebase}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{SECALCUL}
\usepackage{lie}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.50.0.2953}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{BibliographyScheme=Manual}
%TCIDATA{Created=Mon Dec 15 16:20:00 1997}
%TCIDATA{LastRevised=Tuesday, March 25, 2008 11:30:47}
%TCIDATA{<META NAME="ViewSettings" CONTENT="0">}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{CSTFile=SECALCUL.cst}

\input tcilatex
\newenvironment{instructions}{\STARTINSTR}{\ENDINSTR}
\begin{document}


\chapter{1\quad Functions and Models}

\section{1.2\quad Mathematical Models:\ A Catalog of Essential Functions}

%TCIMACRO{\TeXButton{noCCC}{\noCCC}}%
%BeginExpansion
\noCCC%
%EndExpansion
\setET%
%TCIMACRO{\TeXButton{noRM}{\renewcommand{\RM}{0}}}%
%BeginExpansion
\renewcommand{\RM}{0}%
%EndExpansion
%TCIMACRO{\TeXButton{setRM}{\renewcommand{\RM}{1}}}%
%BeginExpansion
\renewcommand{\RM}{1}%
%EndExpansion
%TCIMACRO{%
%\TeXButton{set page 18/ 15}{\ifnum\RM=1 \setcounter{page}{18}\else \setcounter{page}{15}\fi}}%
%BeginExpansion
\ifnum\RM=1 \setcounter{page}{18}\else \setcounter{page}{15}\fi%
%EndExpansion

A \textbf{mathematical model} is a mathematical description (often by means
of a function or an equation) of a real-world phenomenon such as the size of
a population, the demand for a product, the speed of a falling object, the
concentration of a product in a chemical reaction, the life expectancy of a
person at birth, or the cost of emission reductions. The purpose of the
model is to understand the phenomenon and perhaps to make predictions about
future behavior.

Figure 1 illustrates the process of mathematical modeling. Given a
real-world problem, our first task is to formulate a mathematical model by
identifying and naming the independent and dependent variables and making
assumptions that simplify the phenomenon enough to make it mathematically
tractable. We use our knowledge of the physical situation and our
mathematical skills to obtain equations that relate the variables. In
situations where there is no physical law to guide us, we may need to
collect data (either from a library or the Internet or by conducting our own
experiments) and examine the data in the form of a table in order to discern
patterns. From this numerical representation of a function we may wish to
obtain a graphical representation by plotting the data. The graph might even
suggest a suitable algebraic formula in some cases.

%TCIMACRO{\TeXButton{graphicS}{\vspace{12pt}\hskip-190pt\hfil}}%
%BeginExpansion
\vspace{12pt}\hskip-190pt\hfil%
%EndExpansion
\FRAME{itbpFU}{7.1915in}{0.8225in}{0in}{\Qcb{\QTR{FigureNumber}{FIGURE 1}%
\quad The modeling process}}{\Qlb{FIGURE 1 The modeling process}}{%
6et010201_00911.ai}{\special{language "Scientific Word";type
"GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width
7.1915in;height 0.8225in;depth 0in;original-width 7.0638in;original-height
0.7831in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename
'graphics/6et010201_00911.ai';file-properties "XNPEU";}}%
%TCIMACRO{\TeXButton{graphicE}{\vspace{12pt}\hfil}}%
%BeginExpansion
\vspace{12pt}\hfil%
%EndExpansion

The second stage is to apply the mathematics that we know (such as the
calculus that will be developed throughout this book) to the mathematical
model that we have formulated in order to derive mathematical conclusions.
Then, in the third stage, we take those mathematical conclusions and
interpret them as information about the origi\-nal real-world phenomenon by
way of offering explanations or making predictions. The final step is to
test our predictions by checking against new real data. If the predictions
don't compare well with reality, we need to refine our model or to formulate
a new model and start the cycle again.

A mathematical model is never a completely accurate representation of a
physical situation---it is an \textit{idealization}. A good model simplifies
reality enough to permit mathematical calculations but is accurate enough to
provide valuable conclusions. It is important to realize the limitations of
the model. In the end, Mother Nature has the final say.

There are many different types of functions that can be used to model
relationships observed in the real world. In what follows, we discuss the
behavior and graphs of these functions and give examples of situations
appropriately modeled by such functions.%
%TCIMACRO{\TeXButton{longpage}{\enlargethispage{\baselineskip}}}%
%BeginExpansion
\enlargethispage{\baselineskip}%
%EndExpansion

\subsection{LINEAR MODELS}

When we say that $y$ is a \textbf{linear function} of $x$, 
\marginpar{
The coordinate geometry of lines is reviewed in Appendix B.}we mean that the
graph of the function is a line, so we can use the slope-intercept form of
the equation of a line to write a formula for the function as 
\begin{equation*}
y=f(x)=mx+b
\end{equation*}
where $m$ is the slope of the line and $b$ is the $y$-intercept.

A characteristic feature of linear functions is that they grow at a constant
rate. For instance, Figure 2 shows a graph of the linear function $f(x)=3x-2$
and a table of sample values. Notice that whenever \textit{x} increases by
0.1, the value of $f(x)$ increases by 0.3. So $f(x)$ increases three times
as fast as \textit{x}. Thus the slope of the graph $y=3x-2$, namely 3, can
be interpreted as the rate of change of \textit{y} with respect to \textit{x}%
.\bigskip

$\hfill $\FRAME{itbpFU}{136.5pt}{112.5625pt}{48.1875pt}{\Qcb{%
\QTR{FigureNumber}{FIGURE 2}}}{}{3c010202.wmf}{\special{language "Scientific
Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file
"F";width 136.5pt;height 112.5625pt;depth 48.1875pt;original-width
136.5pt;original-height 112.5625pt;cropleft "0";croptop "1";cropright
"1";cropbottom "0";filename 'graphics/3c010202.wmf';file-properties "XNPEU";}%
}\hspace*{12pt}\qquad 
\begin{tabular}{|c|c|}
\hline
\hspace*{12pt}$x$\hspace*{12pt} & \hspace*{12pt}$f(x)=3x-2%
\rule[-6pt]{0pt}{17pt}$\hspace*{12pt} \\ \hline
1.0 & 1.0 \\ 
1.1 & 1.3 \\ 
1.2 & 1.6 \\ 
1.3 & 1.9 \\ 
1.4 & 2.2 \\ 
1.5 & 2.5 \\ \hline
\end{tabular}
$\hfill $


\begin{Solution}
$\,\,\vspace{-4pt}$

\begin{enumerate}
\item[(a)] $f(x)=$ $5^{x}$ is an exponential function. (The $x$ is the
exponent.)

\item[(b)] $g(x)=x^{5}$ is a power function. (The $x$ is the base.) We could
also consider it to be a polynomial of degree 5.

\item[(c)] $h(x)=\dfrac{1+x}{1-\sqrt{x}}$ is an algebraic function.

\item[(d)] $u(t)=1-t+5t^{4}$ is a polynomial of degree 4.$\vspace{-12pt}%
\blacksquare $
\end{enumerate}
\end{Solution}

\QTP{MultColDiv}
Exercises 1.2\vspace{-6pt}

%TCIMACRO{%
%\TeXButton{s2col}{\setlength{\columnsep}{24pt}
%\advance \leftskip by -165pt
%\advance\hsize by 165pt
%\advance\linewidth by 165pt
%\begin{multicols}{2}}}%
%BeginExpansion
\setlength{\columnsep}{24pt}
\advance \leftskip by -165pt
\advance\hsize by 165pt
\advance\linewidth by 165pt
\begin{multicols}{2}%
%EndExpansion

\begin{instructions}
\hspace{9pt}\QTR{SpanExer}{1--2}{\small 
%TCIMACRO{\TeXButton{SQR}{\hskip .5em\rule{4pt}{4pt}\hskip .5em}}%
%BeginExpansion
\hskip .5em\rule{4pt}{4pt}\hskip .5em%
%EndExpansion
Classify each function as a power function, root function, }\newline
{\small polynomial (state its degree), rational function, algebraic
function, }\newline
{\small trigonometric function, exponential function, or logarithmic }%
\newline
{\small function.}
\end{instructions}

\begin{ExerciseList}
\item[$\hfill $1.] 

\begin{ExerciseList}
\item[(a)] $f(x)=\sqrt[5]{x}$\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00067}}}{}{\fbox{\textbf{master 00067}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00067}}}{}{\fbox{\textbf{master 00067}}}{}%
%EndExpansion

\item[(b)] $g(x)=\sqrt{1-x^{2}}$\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00068}}}{}{\fbox{\textbf{master 00068}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00068}}}{}{\fbox{\textbf{master 00068}}}{}%
%EndExpansion

\item[(c)] $h(x)=x^{9}+x^{4}$\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00069}}}{}{\fbox{\textbf{master 00069}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00069}}}{}{\fbox{\textbf{master 00069}}}{}%
%EndExpansion

\item[(d)] $r(x)=\dfrac{x^{2}+1}{x^{3}+x}$\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00070}}}{}{\fbox{\textbf{master 00070}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00070}}}{}{\fbox{\textbf{master 00070}}}{}%
%EndExpansion

\item[(e)] $s(x)=\tan 2x$\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00071}}}{}{\fbox{\textbf{master 00071}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00071}}}{}{\fbox{\textbf{master 00071}}}{}%
%EndExpansion

\item[(f)] $t(x)=\log _{10}x\vspace{3pt}$\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00072}}}{}{\fbox{\textbf{master 00072}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00072}}}{}{\fbox{\textbf{master 00072}}}{}%
%EndExpansion
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~Root\quad (b)~Algebraic\quad (c)~Polynomial (degree 9)\quad (d)~Rational\quad (e) ~Trigonometric\quad (f)~Logarithmic}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~Root\quad (b)~Algebraic\quad (c)~Polynomial (degree 9)\quad (d)~Rational\quad (e) ~Trigonometric\quad (f)~Logarithmic}{}%
%EndExpansion

\item[$\hfill $2.] 

\begin{ExerciseList}
\item[(a)] $y=\dfrac{x-6}{x+6}$\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00073}}}{}{\fbox{\textbf{master 00073}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00073}}}{}{\fbox{\textbf{master 00073}}}{}%
%EndExpansion

\item[(b)] $y=x+\dfrac{x^{2}}{\sqrt{x-1}}$\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00074}}}{}{\fbox{\textbf{master 00074}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00074}}}{}{\fbox{\textbf{master 00074}}}{}%
%EndExpansion

\item[(c)] $y=10^{x}$\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00075}}}{}{\fbox{\textbf{master 00075}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00075}}}{}{\fbox{\textbf{master 00075}}}{}%
%EndExpansion

\item[(d)] $y=x^{10}$\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00076}}}{}{\fbox{\textbf{master 00076}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00076}}}{}{\fbox{\textbf{master 00076}}}{}%
%EndExpansion

\item[(e)] $y=2t^{6}+t^{4}-\pi $\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00077}}}{}{\fbox{\textbf{master 00077}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00077}}}{}{\fbox{\textbf{master 00077}}}{}%
%EndExpansion

\item[(f)] $y=\cos \,\theta +\sin \,\theta $\qquad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00078}}}{}{\fbox{\textbf{master 00078}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00078}}}{}{\fbox{\textbf{master 00078}}}{}%
%EndExpansion
\vspace{-6pt}
\end{ExerciseList}
\end{ExerciseList}

\begin{instructions}
\FRAME{itbpF}{235.6875pt}{5.5pt}{0pt}{}{}{dots.wmf}{\special{language
"Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display
"USEDEF";valid_file "F";width 235.6875pt;height 5.5pt;depth
0pt;original-width 3.9167in;original-height 0.0735in;cropleft "0";croptop
"0.9772";cropright "0.8327";cropbottom "0.0226";filename
'graphics/dots.wmf';file-properties "XNPEU";}}\vspace{-6pt}
\end{instructions}

\begin{instructions}
\QTR{SpanExer}{3--4}{\small 
%TCIMACRO{\TeXButton{SQR}{\hskip .5em\rule{4pt}{4pt}\hskip .5em}}%
%BeginExpansion
\hskip .5em\rule{4pt}{4pt}\hskip .5em%
%EndExpansion
Match each equation with its graph. Explain your choices. (Don't use a
computer or graphing calculator.)}
\end{instructions}

\begin{ExerciseList}
\item[{\hfill \protect\fbox{3.\hspace{-2pt}}}] 

\begin{ExerciseList}
\item[(a)] $y=x^{2}$\hspace{36pt}\textbf{(b)}\hspace{5pt}$y=x^{5}$\hspace{%
36pt}\textbf{(c)}\hspace{5pt}$y=x^{8}$\vspace*{6pt}
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~$h$\quad (b)~$f$\quad (c)~$g$}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~$h$\quad (b)~$f$\quad (c)~$g$}{}%
%EndExpansion

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00079}}}{}{\fbox{\textbf{master 00079}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00079}}}{}{\fbox{\textbf{master 00079}}}{}%
%EndExpansion

\hspace*{\fill}\FRAME{itbpF}{1.9726in}{1.8054in}{0in}{}{}{3c0102x03.wmf}{%
\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio
TRUE;display "USEDEF";valid_file "F";width 1.9726in;height 1.8054in;depth
0in;original-width 1.9441in;original-height 1.7772in;cropleft "0";croptop
"1";cropright "1";cropbottom "0";filename
'graphics/3c0102x03.wmf';file-properties "XNPEU";}}\hspace*{\fill}

\item[\hfill 4.] 

\begin{ExerciseList}
\item[(a)] $y=3x\hspace{15pt}$\textbf{(b)}\hspace{5pt}$y=3^{x}\hspace{15pt}$%
\textbf{(c)}\hspace{5pt}$y=x^{3}\hspace{15pt}$\textbf{(d)}\hspace{5pt}$y=%
\sqrt[3]{x}$\vspace*{6pt}
\end{ExerciseList}

\hspace*{\fill}\FRAME{itbpF}{1.9581in}{1.7771in}{0in}{}{}{3c0102x04.wmf}{%
\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio
TRUE;display "USEDEF";valid_file "F";width 1.9581in;height 1.7771in;depth
0in;original-width 1.9579in;original-height 1.7772in;cropleft "0";croptop
"1";cropright "1";cropbottom "0";filename
'graphics/3c0102x04.wmf';file-properties "XNPEU";}}\hspace*{\fill}

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00080a}}}{}{\fbox{\textbf{master 00080a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00080a}}}{}{\fbox{\textbf{master 00080a}}}{}%
%EndExpansion
\quad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00080a}}}{}{\fbox{\textbf{master 00080a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00080a}}}{}{\fbox{\textbf{master 00080a}}}{}%
%EndExpansion

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00080c}}}{}{\fbox{\textbf{master 00080c}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00080c}}}{}{\fbox{\textbf{master 00080c}}}{}%
%EndExpansion
\quad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00080d}}}{}{\fbox{\textbf{master 00080d}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00080d}}}{}{\fbox{\textbf{master 00080d}}}{}%
%EndExpansion
\vspace{-24pt}
\end{ExerciseList}

\begin{instructions}
\FRAME{itbpF}{235.6875pt}{5.5pt}{0pt}{}{}{dots.wmf}{\special{language
"Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display
"USEDEF";valid_file "F";width 235.6875pt;height 5.5pt;depth
0pt;original-width 3.9167in;original-height 0.0735in;cropleft "0";croptop
"0.9772";cropright "0.8327";cropbottom "0.0226";filename
'graphics/dots.wmf';file-properties "XNPEU";}}
\end{instructions}

\begin{ExerciseList}
\item[{\hfill \protect\fbox{5.\hspace{-2pt}}}] 

\begin{ExerciseList}
\item[(a)] Find an equation for the family of linear functions with slope 2
and sketch several members of the family.

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~$y=2x+b$, where $b$ is the \textit{y}-intercept\newline
%\FRAME{itbpF}{1.2496in}{1.3197in}{1.0032in}{}{}{4a010205a.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.2496in;height 1.3197in;depth 1.0032in;original-width 6.1229in;original-height 6.4627in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename '../../../4c3/4c-SW-ch01/graphics/4a010205a.wmf';file-properties "XNPEU";}}\newline
%}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~$y=2x+b$, where $b$ is the \textit{y}-intercept\newline
\FRAME{itbpF}{1.2496in}{1.3197in}{1.0032in}{}{}{4a010205a.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.2496in;height 1.3197in;depth 1.0032in;original-width 6.1229in;original-height 6.4627in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename '../../../4c3/4c-SW-ch01/graphics/4a010205a.wmf';file-properties "XNPEU";}}\newline
}{}%
%EndExpansion

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00081a}}}{}{\fbox{\textbf{master 00081a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00081a}}}{}{\fbox{\textbf{master 00081a}}}{}%
%EndExpansion

\item[(b)] Find an equation for the family of linear functions such that $%
f(2)=1$ and sketch several members of the family.

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:}\ \FRAME{itbpF}{1.2496in}{1.3197in}{1.0032in}{}{}{4a010205a.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.2496in;height 1.3197in;depth 1.0032in;original-width 6.1229in;original-height 6.4627in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename '../../../4c3/4c-SW-ch01/graphics/4a010205a.wmf';file-properties "XNPEU";}}\newline
%(b)~$y=mx+1-2m$, where $m$ is the slope\newline
%\FRAME{itbpF}{1.2081in}{1.2081in}{1.0032in}{}{}{4a010205b.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.2081in;height 1.2081in;depth 1.0032in;original-width 6.4627in;original-height 6.4627in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename '../../../4c3/4c-SW-ch01/graphics/4a010205b.wmf';file-properties "XNPEU";}}\newline
%}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ \FRAME{itbpF}{1.2496in}{1.3197in}{1.0032in}{}{}{4a010205a.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.2496in;height 1.3197in;depth 1.0032in;original-width 6.1229in;original-height 6.4627in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename '../../../4c3/4c-SW-ch01/graphics/4a010205a.wmf';file-properties "XNPEU";}}\newline
(b)~$y=mx+1-2m$, where $m$ is the slope\newline
\FRAME{itbpF}{1.2081in}{1.2081in}{1.0032in}{}{}{4a010205b.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.2081in;height 1.2081in;depth 1.0032in;original-width 6.4627in;original-height 6.4627in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename '../../../4c3/4c-SW-ch01/graphics/4a010205b.wmf';file-properties "XNPEU";}}\newline
}{}%
%EndExpansion

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00081b}}}{}{\fbox{\textbf{master 00081b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00081b}}}{}{\fbox{\textbf{master 00081b}}}{}%
%EndExpansion

\item[(c)] Which function belongs to both families?

%TCIMACRO{\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (c)~$y=2x-3$}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (c)~$y=2x-3$}{}%
%EndExpansion

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00081c}}}{}{\fbox{\textbf{master 00081c}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00081c}}}{}{\fbox{\textbf{master 00081c}}}{}%
%EndExpansion
\end{ExerciseList}

\item[\hfill 6.] What do all members of the family of linear functions 
\newline
$f(x)=1+m(x+3)$ have in common? Sketch several members of the family.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00082}}}{}{\fbox{\textbf{master 00082}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00082}}}{}{\fbox{\textbf{master 00082}}}{}%
%EndExpansion

\item[\hfill 7.] What do all members of the family of linear functions 
\newline
$f(x)=c-x$ have in common? Sketch several members of the family.

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:} Their graphs have slope $-1$.\newline
%\FRAME{itbpF}{1.5143in}{1.2488in}{0in}{}{}{5et010207.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.5143in;height 1.2488in;depth 0in;original-width 1.5143in;original-height 1.2488in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/5et010207.wmf';file-properties "XNPEU";}}}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:} Their graphs have slope $-1$.\newline
\FRAME{itbpF}{1.5143in}{1.2488in}{0in}{}{}{5et010207.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.5143in;height 1.2488in;depth 0in;original-width 1.5143in;original-height 1.2488in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/5et010207.wmf';file-properties "XNPEU";}}}{}%
%EndExpansion

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00083}}}{}{\fbox{\textbf{master 00083}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00083}}}{}{\fbox{\textbf{master 00083}}}{}%
%EndExpansion

\item[\hfill 8.] Find expressions for the quadratic functions whose graphs
are shown.

\FRAME{dtbpF}{2.7492in}{0.9816in}{0pt}{}{}{6et0102x08_00076.ai}{\special%
{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio
TRUE;display "USEDEF";valid_file "F";width 2.7492in;height 0.9816in;depth
0pt;original-width 2.7224in;original-height 0.9539in;cropleft "0";croptop
"1";cropright "1";cropbottom "0";filename
'graphics/6et0102x08_00076.ai';file-properties "XNPEU";}}

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 30016}}}{}{\fbox{\textbf{master 30016}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 30016}}}{}{\fbox{\textbf{master 30016}}}{}%
%EndExpansion

\item[9.] Find an expression for a cubic function $f$ if $f(1)=6$ and

$f(-1)=f(0)=f(2)=0$.

%TCIMACRO{\hyperref{ANSWER}{}{\textbf{ANSWER:} $f(x)=-3x(x+1)(x-2)$}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:} $f(x)=-3x(x+1)(x-2)$}{}%
%EndExpansion

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 30017}}}{}{\fbox{\textbf{master 30017}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 30017}}}{}{\fbox{\textbf{master 30017}}}{}%
%EndExpansion

\item[10.] Recent studies indicate that the average surface temperature of
the earth has been rising steadily. Some scientists have modeled the
temperature by the linear function $T=0.02t+8.50$, where $T$ is temperature
in $^{\circ }\mathrm{C}$ and $t$ represents years since 1900.

\begin{ExerciseList}
\item[(a)] What do the slope and $T$-intercept represent?
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 30018a}}}{}{\fbox{\textbf{master 30018a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 30018a}}}{}{\fbox{\textbf{master 30018a}}}{}%
%EndExpansion

\begin{ExerciseList}
\item[(b)] Use the equation to predict the average global surface
temperature in 2100.
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 30018b}}}{}{\fbox{\textbf{master 30018b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 30018b}}}{}{\fbox{\textbf{master 30018b}}}{}%
%EndExpansion

\item[11.] If the recommended adult dosage for a drug is $D$ (in mg), then
to determine the appropriate dosage $c$ for a child of age $a$, pharmacists
use the equation $c=0.0417D(a+1)$. Suppose the dosage for an adult is 200 mg.

\begin{ExerciseList}
\item[(a)] Find the slope of the graph of $c$. What does it represent?
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 30019a}}}{}{\fbox{\textbf{master 30019a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 30019a}}}{}{\fbox{\textbf{master 30019a}}}{}%
%EndExpansion

\begin{ExerciseList}
\item[(b)] What is the dosage for a newborn?
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 30019b}}}{}{\fbox{\textbf{master 30019b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 30019b}}}{}{\fbox{\textbf{master 30019b}}}{}%
%EndExpansion

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:} (a)~8.34, change in mg for every 1 year change\quad (b)~8.34 mg}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:} (a)~8.34, change in mg for every 1 year change\quad (b)~8.34 mg}{}%
%EndExpansion

\item[12.] The manager of a weekend flea market knows from past experience
that if he charges $x$ dollars for a rental space at the flea market, then
the number $y$ of spaces he can rent is given by the equation $y=200-4x$.%
\vspace{-2pt}

\begin{ExerciseList}
\item[(a)] Sketch a graph of this linear function. (Remember that the rental
charge per space and the number of spaces rented can't be negative
quantities.)

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00084a}}}{}{\fbox{\textbf{master 00084a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00084a}}}{}{\fbox{\textbf{master 00084a}}}{}%
%EndExpansion

\item[(b)] What do the slope, the $y$-intercept, and the $x$-intercept of
the graph represent?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00084b}}}{}{\fbox{\textbf{master 00084b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00084b}}}{}{\fbox{\textbf{master 00084b}}}{}%
%EndExpansion
\end{ExerciseList}

\item[13.] The relationship between the Fahrenheit ($F$) and Celsius ($C$)
temperature scales is given by the linear function \newline
$F=\frac{9}{5}C+32$.\vspace{-2pt}

\begin{ExerciseList}
\item[(a)] Sketch a graph of this function.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00085a}}}{}{\fbox{\textbf{master 00085a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00085a}}}{}{\fbox{\textbf{master 00085a}}}{}%
%EndExpansion

\item[(b)] What is the slope of the graph and what does it represent? What
is the $F$-intercept and what does it represent?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00085b}}}{}{\fbox{\textbf{master 00085b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00085b}}}{}{\fbox{\textbf{master 00085b}}}{}%
%EndExpansion
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~\FRAME{itbpF}{1.4382in}{1.2981in}{1.254in}{}{}{4a010207a.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.4382in;height 1.2981in;depth 1.254in;original-width 6.4627in;original-height 5.8237in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/4a010207a.wmf';file-properties "XNPEU";}}\newline
%(b)~$\tfrac{9}{5}$, change in $^{\circ }$F for every 1$^{\circ }$C change; 32, Fahrenheit temperature corresponding to 0$^{\circ }$C}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~\FRAME{itbpF}{1.4382in}{1.2981in}{1.254in}{}{}{4a010207a.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.4382in;height 1.2981in;depth 1.254in;original-width 6.4627in;original-height 5.8237in;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/4a010207a.wmf';file-properties "XNPEU";}}\newline
(b)~$\tfrac{9}{5}$, change in $^{\circ }$F for every 1$^{\circ }$C change; 32, Fahrenheit temperature corresponding to 0$^{\circ }$C}{}%
%EndExpansion

\item[14.] Jason leaves Detroit at 2:00 \textsc{pm} and drives at a constant
speed west along I-96. He passes Ann Arbor, 40 mi from Detroit, at 2:50 
\textsc{pm}.\textsc{\vspace{-2pt}}

\begin{ExerciseList}
\item[(a)] Express the distance traveled in terms of the time elapsed.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00086a}}}{}{\fbox{\textbf{master 00086a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00086a}}}{}{\fbox{\textbf{master 00086a}}}{}%
%EndExpansion

\item[(b)] Draw the graph of the equation in part (a).

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00086b}}}{}{\fbox{\textbf{master 00086b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00086b}}}{}{\fbox{\textbf{master 00086b}}}{}%
%EndExpansion

\item[(c)] What is the slope of this line? What does it represent?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00086c}}}{}{\fbox{\textbf{master 00086c}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00086c}}}{}{\fbox{\textbf{master 00086c}}}{}%
%EndExpansion
\end{ExerciseList}

\item[{\hfill \protect\fbox{\hspace{-2pt}15.\hspace{-2pt}}}] Biologists have
noticed that the chirping rate of crickets of a certain species is related
to temperature, and the relationship appears to be very nearly linear. A
cricket produces 113 chirps per minute at $70^{\circ }$F and 173 chirps per
minute at $80^{\circ }$F.\vspace{-2pt}

\begin{ExerciseList}
\item[(a)] Find a linear equation that models the temperature $T$ as a
function of the number of chirps per minute $N$.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00087a}}}{}{\fbox{\textbf{master 00087a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00087a}}}{}{\fbox{\textbf{master 00087a}}}{}%
%EndExpansion

\item[(b)] What is the slope of the graph? What does it represent?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00087b}}}{}{\fbox{\textbf{master 00087b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00087b}}}{}{\fbox{\textbf{master 00087b}}}{}%
%EndExpansion

\item[(c)] If the crickets are chirping at 150 chirps per minute, estimate
the temperature.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00087c}}}{}{\fbox{\textbf{master 00087c}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00087c}}}{}{\fbox{\textbf{master 00087c}}}{}%
%EndExpansion
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~$T=\tfrac{1}{6}N+\tfrac{307}{6}$\quad (b)~$\tfrac{1}{6}$, change in $^{\circ }$F for every chirp per minute change\quad (c)~76$^{\circ }$F}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~$T=\tfrac{1}{6}N+\tfrac{307}{6}$\quad (b)~$\tfrac{1}{6}$, change in $^{\circ }$F for every chirp per minute change\quad (c)~76$^{\circ }$F}{}%
%EndExpansion

\item[16.] The manager of a furniture factory finds that it costs \$2200 to
manufacture 100 chairs in one day and \$4800 to produce 300 chairs in one
day.\vspace{-2pt}

\begin{ExerciseList}
\item[(a)] Express the cost as a function of the number of chairs produced,
assuming that it is linear. Then sketch the graph.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00088a}}}{}{\fbox{\textbf{master 00088a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00088a}}}{}{\fbox{\textbf{master 00088a}}}{}%
%EndExpansion

\item[(b)] What is the slope of the graph and what does it represent?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00088b}}}{}{\fbox{\textbf{master 00088b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00088b}}}{}{\fbox{\textbf{master 00088b}}}{}%
%EndExpansion

\item[(c)] What is the $y$-intercept of the graph and what does it \newline
represent?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00088c}}}{}{\fbox{\textbf{master 00088c}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00088c}}}{}{\fbox{\textbf{master 00088c}}}{}%
%EndExpansion
\end{ExerciseList}

\item[{\hfill \protect\fbox{\hspace{-2pt}17.\hspace{-2pt}}}] At the surface
of the ocean, the water pressure is the same as the air pressure above the
water, 15 lb$/$in$^{2}$. Below the surface, the water pressure increases by
4.34 lb$/$in$^{2}$ for every 10 ft of descent.\vspace{-2pt}

\begin{ExerciseList}
\item[(a)] Express the water pressure as a function of the depth below the
ocean surface.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00089a}}}{}{\fbox{\textbf{master 00089a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00089a}}}{}{\fbox{\textbf{master 00089a}}}{}%
%EndExpansion

\item[(b)] At what depth is the pressure 100 lb$/$in$^{2}$?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00089b}}}{}{\fbox{\textbf{master 00089b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00089b}}}{}{\fbox{\textbf{master 00089b}}}{}%
%EndExpansion
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~$P=0.434d+15$\quad (b)~196~ft}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~$P=0.434d+15$\quad (b)~196~ft}{}%
%EndExpansion

\item[18.] The monthly cost of driving a car depends on the number of miles
driven. Lynn found that in May it cost her \$380 to drive 480 mi and in June
it cost her \$460 to drive 800 mi.\vspace{-2pt}

\begin{ExerciseList}
\item[(a)] Express the monthly cost $C$ as a function of the distance driven 
$d$, assuming that a linear relationship gives a suitable model.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00090a}}}{}{\fbox{\textbf{master 00090a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00090a}}}{}{\fbox{\textbf{master 00090a}}}{}%
%EndExpansion

\item[(b)] Use part (a) to predict the cost of driving 1500 miles per month.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00090b}}}{}{\fbox{\textbf{master 00090b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00090b}}}{}{\fbox{\textbf{master 00090b}}}{}%
%EndExpansion

\item[(c)] Draw the graph of the linear function. What does the slope
represent?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00090c}}}{}{\fbox{\textbf{master 00090c}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00090c}}}{}{\fbox{\textbf{master 00090c}}}{}%
%EndExpansion

\item[(d)] What does the $y$-intercept represent?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00090d}}}{}{\fbox{\textbf{master 00090d}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00090d}}}{}{\fbox{\textbf{master 00090d}}}{}%
%EndExpansion

\item[(e)] Why does a linear function give a suitable model in this
situation?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00090e}}}{}{\fbox{\textbf{master 00090e}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00090e}}}{}{\fbox{\textbf{master 00090e}}}{}%
%EndExpansion
\end{ExerciseList}
\end{ExerciseList}

\begin{instructions}
\QTR{SpanExer}{19--20}{\small 
%TCIMACRO{\TeXButton{SQR}{\hskip .5em\rule{4pt}{4pt}\hskip .5em}}%
%BeginExpansion
\hskip .5em\rule{4pt}{4pt}\hskip .5em%
%EndExpansion
For each scatter plot, decide what type of function you might choose as a
model for the data. Explain your choices.}
\end{instructions}

\begin{ExerciseList}
\item[19.] 

\begin{ExerciseList}
\item[(a)] \ \FRAME{itbpF}{1.4304in}{1.4304in}{1.305in}{}{}{4e0102x13a.wmf}{%
\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio
TRUE;display "USEDEF";valid_file "F";width 1.4304in;height 1.4304in;depth
1.305in;original-width 1.4304in;original-height 1.4304in;cropleft
"0";croptop "1";cropright "1";cropbottom "0";filename
'graphics/4e0102x13a.wmf';file-properties "XNPEU";}}\quad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00091}}}{}{\fbox{\textbf{master 00091}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00091}}}{}{\fbox{\textbf{master 00091}}}{}%
%EndExpansion
\vspace{6pt}

%TCIMACRO{\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~Cosine}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~Cosine}{}%
%EndExpansion

\item[(b)] \QTR{PartLetter}{\ }\FRAME{itbpF}{1.4304in}{1.4304in}{1.305in}{}{%
}{4e0102x13b.wmf}{\special{language "Scientific Word";type
"GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width
1.4304in;height 1.4304in;depth 1.305in;original-width
1.4304in;original-height 1.4304in;cropleft "0";croptop "1";cropright
"1";cropbottom "0";filename 'graphics/4e0102x13b.wmf';file-properties
"XNPEU";}}\quad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00092}}}{}{\fbox{\textbf{master 00092}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00092}}}{}{\fbox{\textbf{master 00092}}}{}%
%EndExpansion

%TCIMACRO{\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (b)~Linear}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (b)~Linear}{}%
%EndExpansion
\end{ExerciseList}

\item[20.] 

\begin{ExerciseList}
\item[(a)] \ \FRAME{itbpF}{1.2635in}{1.2782in}{1.305in}{}{}{4e0102x14a.wmf}{%
\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio
TRUE;display "USEDEF";valid_file "F";width 1.2635in;height 1.2782in;depth
1.305in;original-width 1.2635in;original-height 1.2782in;cropleft
"0";croptop "1";cropright "1";cropbottom "0";filename
'graphics/4e0102x14a.wmf';file-properties "XNPEU";}}\quad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00093}}}{}{\fbox{\textbf{master 00093}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00093}}}{}{\fbox{\textbf{master 00093}}}{}%
%EndExpansion
\vspace{6pt}

\item[(b)] \QTR{PartLetter}{\ }\FRAME{itbpF}{1.299in}{1.3128in}{1.305in}{}{}{%
4e0102x14b.wmf}{\special{language "Scientific Word";type
"GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width
1.299in;height 1.3128in;depth 1.305in;original-width
1.2635in;original-height 1.2782in;cropleft "0";croptop "1";cropright
"1";cropbottom "0";filename 'graphics/4e0102x14b.wmf';file-properties
"XNPEU";}}\quad 
%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00094}}}{}{\fbox{\textbf{master 00094}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00094}}}{}{\fbox{\textbf{master 00094}}}{}%
%EndExpansion
\vspace{6pt}
\end{ExerciseList}
\end{ExerciseList}

\begin{instructions}
\FRAME{itbpF}{235.6875pt}{5.5pt}{0pt}{}{}{dots.wmf}{\special{language
"Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display
"USEDEF";valid_file "F";width 235.6875pt;height 5.5pt;depth
0pt;original-width 3.9167in;original-height 0.0735in;cropleft "0";croptop
"0.9772";cropright "0.8327";cropbottom "0.0226";filename
'graphics/dots.wmf';file-properties "XNPEU";}}
\end{instructions}

\begin{ExerciseList}
\item[21.] 
%TCIMACRO{\TeXButton{GCALCX}{\GCALCX}}%
%BeginExpansion
\GCALCX%
%EndExpansion
The table shows (lifetime) peptic ulcer rates (per 100 population) for
various family incomes as reported by the National Health Interview
Survey.\bigskip

$\hfill ${\small $%
\begin{tabular}{|c|c|}
\hline
& Ulcer rate \\ 
Income & (per 100 population) \\ \hline
\$\QTR{white}{\hspace*{4pt}}4,000 & 14.1 \\ 
\$\QTR{white}{\hspace*{4pt}}6,000 & 13.0 \\ 
\$\QTR{white}{\hspace*{4pt}}8,000 & 13.4 \\ 
\$12,000 & 12.5 \\ 
\$16,000 & 12.0 \\ 
\$20,000 & 12.4 \\ 
\$30,000 & 10.5 \\ 
\$45,000 & \QTR{white}{\hspace*{4pt}}9.4 \\ 
\$60,000 & \QTR{white}{\hspace*{4pt}}8.2 \\ \hline
\end{tabular}%
\ $}$\hfill $\bigskip

\begin{ExerciseList}
\item[(a)] Make a scatter plot of these data and decide whether a linear
model is appropriate.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00095a}}}{}{\fbox{\textbf{master 00095a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00095a}}}{}{\fbox{\textbf{master 00095a}}}{}%
%EndExpansion

\item[(b)] Find and graph a linear model using the first and last data
points.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00095b}}}{}{\fbox{\textbf{master 00095b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00095b}}}{}{\fbox{\textbf{master 00095b}}}{}%
%EndExpansion

\item[(c)] Find and graph the least squares regression line.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00095c}}}{}{\fbox{\textbf{master 00095c}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00095c}}}{}{\fbox{\textbf{master 00095c}}}{}%
%EndExpansion

\item[(d)] Use the linear model in part (c) to estimate the ulcer rate for
an income of \$25,000.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00095d}}}{}{\fbox{\textbf{master 00095d}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00095d}}}{}{\fbox{\textbf{master 00095d}}}{}%
%EndExpansion

\item[(e)] According to the model, how likely is someone with an income of
\$80,000 to suffer from peptic ulcers?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00095e}}}{}{\fbox{\textbf{master 00095e}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00095e}}}{}{\fbox{\textbf{master 00095e}}}{}%
%EndExpansion

\item[(f)] Do you think it would be reasonable to apply the model to someone
with an income of \$200,000?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00095f}}}{}{\fbox{\textbf{master 00095f}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00095f}}}{}{\fbox{\textbf{master 00095f}}}{}%
%EndExpansion
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~\FRAME{itbpF}{1.5696in}{0.9842in}{0.8786in}{}{}{4a010215a.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.5696in;height 0.9842in;depth 0.8786in;original-width 0pt;original-height 0pt;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/4a010215a.wmf';file-properties "XNPEU";}}\newline
%Linear model is appropriate.\newline
%(b)~$y=-0.000105x+14.521$\newline
%\FRAME{itbpF}{1.5696in}{0.9842in}{0.8536in}{}{}{4a010215b.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.5696in;height 0.9842in;depth 0.8536in;original-width 0pt;original-height 0pt;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/4a010215b.wmf';file-properties "XNPEU";}}\newline
%(c)~$y=-0.00009979x+13.951$ [See graph in (b).]\quad \newline
%(d)~About 11.5 per 100 population\quad (e)~About 6\%\quad (f)~No}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~\FRAME{itbpF}{1.5696in}{0.9842in}{0.8786in}{}{}{4a010215a.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.5696in;height 0.9842in;depth 0.8786in;original-width 0pt;original-height 0pt;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/4a010215a.wmf';file-properties "XNPEU";}}\newline
Linear model is appropriate.\newline
(b)~$y=-0.000105x+14.521$\newline
\FRAME{itbpF}{1.5696in}{0.9842in}{0.8536in}{}{}{4a010215b.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.5696in;height 0.9842in;depth 0.8536in;original-width 0pt;original-height 0pt;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/4a010215b.wmf';file-properties "XNPEU";}}\newline
(c)~$y=-0.00009979x+13.951$ [See graph in (b).]\quad \newline
(d)~About 11.5 per 100 population\quad (e)~About 6\%\quad (f)~No}{}%
%EndExpansion

\item[22.] 
%TCIMACRO{\TeXButton{GCALCX}{\GCALCX}}%
%BeginExpansion
\GCALCX%
%EndExpansion
Biologists have observed that the chirping rate of crickets of a certain
species appears to be related to temperature. The table shows the chirping
rates for various temperatures.\bigskip

$\hfill ${\small $%
\begin{tabular}{|c|c||c|c|}
\hline
\hspace*{-7pt}%
\begin{tabular}[t]{c}
Temperature \\ 
($^{\circ }$F$)$%
\end{tabular}%
\hspace*{-7pt} & \hspace*{-7pt}%
\begin{tabular}[t]{c}
Chirping rate \\ 
(chirps/min$)$%
\end{tabular}%
\hspace*{-7pt} & \hspace*{-7pt}%
\begin{tabular}[t]{c}
Temperature \\ 
($^{\circ }$F$)$%
\end{tabular}%
\hspace*{-7pt} & \hspace*{-7pt}%
\begin{tabular}[t]{c}
Chirping rate \\ 
(chirps/min$)$%
\end{tabular}%
\hspace*{-7pt} \\ \hline
50 & 20 & 75 & 140 \\ 
55 & 46 & 80 & 173 \\ 
60 & 79 & 85 & 198 \\ 
65 & 91 & 90 & 211 \\ 
70 & 113 &  &  \\ \hline
\end{tabular}%
\ \ $}$\hfill $\bigskip

\begin{ExerciseList}
\item[(a)] Make a scatter plot of the data.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00096a}}}{}{\fbox{\textbf{master 00096a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00096a}}}{}{\fbox{\textbf{master 00096a}}}{}%
%EndExpansion

\item[(b)] Find and graph the regression line.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00096b}}}{}{\fbox{\textbf{master 00096b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00096b}}}{}{\fbox{\textbf{master 00096b}}}{}%
%EndExpansion

\item[(c)] Use the linear model in part (b) to estimate the chirping rate at 
$100^{\circ }$F.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00096c}}}{}{\fbox{\textbf{master 00096c}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00096c}}}{}{\fbox{\textbf{master 00096c}}}{}%
%EndExpansion
\end{ExerciseList}

\item[23.] 
%TCIMACRO{\TeXButton{GCALCX}{\GCALCX}}%
%BeginExpansion
\GCALCX%
%EndExpansion
The table gives the winning heights for the Olympic pole vault competitions
in the 20th century.\bigskip

$\hfill ${\small $%
\begin{tabular}{|c|c||c|c|}
\hline
Year & Height (ft) & Year & Height (ft) \\ \hline
1900 & 10.83 & 1956 & 14.96 \\ 
1904 & 11.48 & 1960 & 15.42 \\ 
1908 & 12.17 & 1964 & 16.73 \\ 
1912 & 12.96 & 1968 & 17.71 \\ 
1920 & 13.42 & 1972 & 18.04 \\ 
1924 & 12.96 & 1976 & 18.04 \\ 
1928 & 13.77 & 1980 & 18.96 \\ 
1932 & 14.15 & 1984 & 18.85 \\ 
1936 & 14.27 & 1988 & 19.77 \\ 
1948 & 14.10 & 1992 & 19.02 \\ 
1952 & 14.92 & 1996 & 19.42 \\ \hline
\end{tabular}%
$}$\hfill $\bigskip

\begin{ExerciseList}
\item[(a)] Make a scatter plot and decide whether a linear model is
appropriate.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00097a}}}{}{\fbox{\textbf{master 00097a}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00097a}}}{}{\fbox{\textbf{master 00097a}}}{}%
%EndExpansion

\item[(b)] Find and graph the regression line.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00097b}}}{}{\fbox{\textbf{master 00097b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00097b}}}{}{\fbox{\textbf{master 00097b}}}{}%
%EndExpansion

\item[(c)] Use the linear model to predict the height of the winning pole
vault at the 2000 Olympics and compare with the actual winning height of
19.36 feet.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00097c}}}{}{\fbox{\textbf{master 00097c}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00097c}}}{}{\fbox{\textbf{master 00097c}}}{}%
%EndExpansion

\item[(d)] Is it reasonable to use the model to predict the winning height
at the 2100 Olympics?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00097d}}}{}{\fbox{\textbf{master 00097d}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00097d}}}{}{\fbox{\textbf{master 00097d}}}{}%
%EndExpansion
\end{ExerciseList}

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~\FRAME{itbpF}{1.8749in}{1.011in}{1.0032in}{}{}{4a010217a.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.8749in;height 1.011in;depth 1.0032in;original-width 0pt;original-height 0pt;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/4a010217a.wmf';file-properties "XNPEU";}}\newline
%Linear model is appropriate.\newline
%(b)~$y=0.08912x-158.24$\newline
%\FRAME{itbpF}{1.8749in}{1.011in}{1.0032in}{}{}{4a010217b.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.8749in;height 1.011in;depth 1.0032in;original-width 0pt;original-height 0pt;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/4a010217b.wmf';file-properties "XNPEU";}}\newline
%(c)~20~ft\quad (d)~No}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\ (a)~\FRAME{itbpF}{1.8749in}{1.011in}{1.0032in}{}{}{4a010217a.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.8749in;height 1.011in;depth 1.0032in;original-width 0pt;original-height 0pt;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/4a010217a.wmf';file-properties "XNPEU";}}\newline
Linear model is appropriate.\newline
(b)~$y=0.08912x-158.24$\newline
\FRAME{itbpF}{1.8749in}{1.011in}{1.0032in}{}{}{4a010217b.wmf}{\special{language "Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display "USEDEF";valid_file "F";width 1.8749in;height 1.011in;depth 1.0032in;original-width 0pt;original-height 0pt;cropleft "0";croptop "1";cropright "1";cropbottom "0";filename 'graphics/4a010217b.wmf';file-properties "XNPEU";}}\newline
(c)~20~ft\quad (d)~No}{}%
%EndExpansion

\item[24.] 
%TCIMACRO{\TeXButton{GCALCX}{\GCALCX}}%
%BeginExpansion
\GCALCX%
%EndExpansion
A study by the US Office of Science and Technology in 1972 estimated the
cost (in 1972 dollars) to reduce automobile emissions by certain
percentages:\bigskip

$\hfill ${\small $%
\begin{tabular}{|c|c||c|c|}
\hline
\hspace*{-7pt}%
\begin{tabular}[t]{c}
Reduction in \\ 
emissions (\%)%
\end{tabular}%
\hspace*{-7pt} & \hspace*{-7pt}%
\begin{tabular}[t]{c}
Cost per \\ 
car (in \$)%
\end{tabular}%
\hspace*{-7pt} & \hspace*{-7pt}%
\begin{tabular}[t]{c}
Reduction in \\ 
emissions (\%)%
\end{tabular}%
\hspace*{-7pt} & \hspace*{-7pt}%
\begin{tabular}[t]{c}
Cost per \\ 
car (in \$)%
\end{tabular}%
\hspace*{-7pt} \\ \hline
50 & 45 & 75 & \hspace{3pt}90 \\ 
55 & 55 & 80 & 100 \\ 
60 & 62 & 85 & 200 \\ 
65 & 70 & 90 & 375 \\ 
70 & 80 & 95 & 600 \\ \hline
\end{tabular}%
\ $}$\hfill $\bigskip

Find a model that captures the \textquotedblleft diminishing
returns\textquotedblright\ trend of these data.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00098}}}{}{\fbox{\textbf{master 00098}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00098}}}{}{\fbox{\textbf{master 00098}}}{}%
%EndExpansion

\item[25.] 
%TCIMACRO{\TeXButton{GCALCX}{\GCALCX}}%
%BeginExpansion
\GCALCX%
%EndExpansion
Use the data in the table to model the population of the world in the 20th
century by a cubic function. Then use your model to estimate the population
in the year 1925.\bigskip

$\hfill ${\small $%
\begin{tabular}{|c|c||c|c|}
\hline
\begin{tabular}{l}
\\ 
Year%
\end{tabular}
& 
\begin{tabular}{l}
Population \\ 
(millions)%
\end{tabular}
& 
\begin{tabular}{l}
\\ 
Year%
\end{tabular}
& 
\begin{tabular}{l}
Population \\ 
(millions)%
\end{tabular}
\\ \hline
1900 & 1650 & 1960 & 3040 \\ 
1910 & 1750 & 1970 & 3710 \\ 
1920 & 1860 & 1980 & 4450 \\ 
1930 & 2070 & 1990 & 5280 \\ 
1940 & 2300 & 2000 & 6080 \\ 
1950 & 2560 &  &  \\ \hline
\end{tabular}%
\ \ \ $}$\hfill $\vspace{3pt}

%TCIMACRO{%
%\hyperref{ANSWER}{}{\textbf{ANSWER:}\newline
%\ $y=0.0012937x^{3}-7.06142x^{2}+12$,$823x-7$,$743$,$770$; \newline
%\textbf{Answer:}\ 1914 million}{}}%
%BeginExpansion
\msihyperref{ANSWER}{}{\textbf{ANSWER:}\newline
\ $y=0.0012937x^{3}-7.06142x^{2}+12$,$823x-7$,$743$,$770$; \newline
\textbf{Answer:}\ 1914 million}{}%
%EndExpansion

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00099}}}{}{\fbox{\textbf{master 00099}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00099}}}{}{\fbox{\textbf{master 00099}}}{}%
%EndExpansion

\item[26.] 
%TCIMACRO{\TeXButton{GCALCX}{\GCALCX}}%
%BeginExpansion
\GCALCX%
%EndExpansion
The table shows the mean (average) distances $d$ of the planets from the sun
(taking the unit of measurement to be the distance from earth to the sun)
and their periods $T$ (time of revolution in years).\bigskip

$\hfill ${\small $%
\begin{tabular}{|c|c|c|}
\hline
Planet & $d$ & $T$ \\ \hline
\multicolumn{1}{|l|}{Mercury} & \multicolumn{1}{|r|}{0.387} & 
\multicolumn{1}{|r|}{0.241} \\ 
\multicolumn{1}{|l|}{Venus} & \multicolumn{1}{|r|}{0.723} & 
\multicolumn{1}{|r|}{0.615} \\ 
\multicolumn{1}{|l|}{Earth} & \multicolumn{1}{|r|}{1.000} & 
\multicolumn{1}{|r|}{1.000} \\ 
\multicolumn{1}{|l|}{Mars} & \multicolumn{1}{|r|}{1.523} & 
\multicolumn{1}{|r|}{1.881} \\ 
\multicolumn{1}{|l|}{Jupiter} & \multicolumn{1}{|r|}{5.203} & 
\multicolumn{1}{|r|}{11.861} \\ 
\multicolumn{1}{|l|}{Saturn} & \multicolumn{1}{|r|}{9.541} & 
\multicolumn{1}{|r|}{29.457} \\ 
\multicolumn{1}{|l|}{Uranus} & \multicolumn{1}{|r|}{19.190} & 
\multicolumn{1}{|r|}{84.008} \\ 
\multicolumn{1}{|l|}{Neptune} & \multicolumn{1}{|r|}{30.086} & 
\multicolumn{1}{|r|}{164.784} \\ \hline
\end{tabular}%
\ \ $}$\hfill $\bigskip

\begin{ExerciseList}
\item[(a)] Fit a power model to the data.

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 81407}}}{}{\fbox{\textbf{master 81407}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 81407}}}{}{\fbox{\textbf{master 81407}}}{}%
%EndExpansion

\item[(b)] Kepler's Third Law of Planetary Motion states that \\[3pt]
\hspace*{8pt}%
\begin{tabular}[t]{p{175pt}}
%TCIMACRO{\TeXButton{rr}{\raggedright}}%
%BeginExpansion
\raggedright%
%EndExpansion
\textquotedblleft The square of the period of revolution of a planet is
proportional to the cube of its mean distance from the sun.\textquotedblright%
\end{tabular}%
\\[3pt]
Does your model corroborate Kepler's Third Law?

%TCIMACRO{%
%\hyperref{\fbox{\textbf{master 00100b}}}{}{\fbox{\textbf{master 00100b}}}{}}%
%BeginExpansion
\msihyperref{\fbox{\textbf{master 00100b}}}{}{\fbox{\textbf{master 00100b}}}{}%
%EndExpansion
\end{ExerciseList}
\end{ExerciseList}

%TCIMACRO{%
%\TeXButton{e2col}{\end{multicols}
%\advance \leftskip by 165pt
%\advance\hsize by -165pt
%\advance\linewidth by -165pt
%}}%
%BeginExpansion
\end{multicols}
\advance \leftskip by 165pt
\advance\hsize by -165pt
\advance\linewidth by -165pt
%
%EndExpansion

\end{document}
